## Factor or Continuous Variables

---
output: html_document
---

Factor Variables = Categorial Variables
Gender = M/F, state = 1-50 + DC, etc
This is how to deal w factor variables

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Open the reddit data used in EG

```{r}
setwd("C:/R/udacity_course")
reddit <-read.csv('reddit.csv')
```

Look at data to see what's there
str yeah k but output is fugly
Can get summary of all w...summary

```{r}
summary(reddit)
```

Look at gender - is 0/1, summary is awful for showing this
Care about 1 - gender 
Table it. Oh look much better. 

```{r}
table(reddit$gender)
```

Now look at Ordered Factors - ie ranked
age.range is a good one for this

```{r}
levels(reddit$age.range)
```

Let's see what it looks like visually.
Load ggplot2 if necessary
Look at output, notice it's not in order
b/c numbers are set first then letters

WHICH IS WHY WE NEED AN ORDERED FACTOR

```{r}
qplot(data=reddit,x=age.range)
```

```{r}
age.range.order <- ordered(reddit$age.range, levels = c("Under 18","18-24","25-34","45-54","65 or Above","NA"))
```

Now notice how this new plot looks better

```{r}
qplot(data=reddit, x=age.range.order)
```

Same w income just do it more programmy wo creating sep value
Looks fugly, get levels to list properly

```{r}
qplot(data=reddit, x=income.range)
levels(reddit$income.range)
```

This tells R to recode the income.range var as a Factor, then to Order it, 
rather than creating some separate value/variable. Looks cleaner to me.

```{r}
reddit$income.range <- factor(reddit$income.range, levels = c("Under $20,000", "$20,000 - $29,999", "$30,000 - $39,999", "$40,000 - $49,999", "$50,000 - $69,999", "$70,000 - $99,999", "$100,000 - $149,999", "$150,000 or more","NA"), ordered=T)
qplot(data=reddit, x=income.range)
```



